Neural Networks for Neural Tumors

### 2025-06-01 latex/draft.tex
renamed
minor adjustments so far

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\graphicspath{{../images/}}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
% \setlength{\parindent}{1em}
% \setlength{\parskip}{0.8em}
\begin{document}

@@

\author{
\IEEEauthorblockN{Joe Vogel}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
jvogel@ucdavis.edu}

\and
\IEEEauthorblockN{Jonathan Mo}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
jtnmo@ucdavis.edu}

\and
\IEEEauthorblockN{Nam Nguyen}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
ncnn@ucdavis.edu}

\and
\IEEEauthorblockN{Aryan Karnwal}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
akarnwal@ucdavis.edu}

\and
\IEEEauthorblockN{Richard Ho}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
rdho@ucdavis.edu}
}

@@

\begin{abstract}
This report documents the process of making a machine learning model to classify brain tumors using MRI brain scan imagery. This project saw the creation of three successful models using the following techniques: CNN, SVM, and Random Forest. 
\end{abstract}

\begin{IEEEkeywords}
Brain tumor classification, Convolutional Neural Network (CNN), MRI brain imagery, Random Forest, Support Vector Machine (SVM), 
\end{IEEEkeywords}

\section{Introduction (1-2 pages)}
Hundreds of thousands of MRI brain scans are performed every day. With MRI scans being so intricate, it can sometimes be hard for even a trained professional to determine whether or not a patient has a brain tumor, and even harder to tell which kind.

\subsection{Motivation}

Having a tool that could detect and identify tumors within seconds, and with high accuracy, would save a lot of time and money. Not to mention the impact it would have on developing nations who might not have as many specially trained doctors that can readily perform MRI analysis. Early detection is absolutely key to diagnosing and treating these tumors.

\subsection{Approach}

A classification machine learning model is the perfect solution to this problem. Using a convolutional neural network, we can train a model on MRI images to be able to accurately classify MRI scans into four classifications: glioma tumor, meningioma tumor, pituitary tumor, and no tumor. With its complex structure, such a model can detect nuances in the scans that are otherwise unnoticeable to the human eye. Given any image of an MRI brain scan, the model should be able to predict these four classes with over 95\% accuracy.

\section{Literature Review (1/2 page)}

\subsection{Brain Tumor Classification in MRI Image Using Convolutional Neural Network}

@@

\subsection{A Deep CNN Based Multi-Class Classification of Alzheimer's Disease Using MRI}

@@

\subsection{Brain Tumor Classification Using Convolutional Neural Networks}

@@

\section{Dataset Description (1 page)}

The dataset used for this project comes from a group of brilliant students at Savitribai Phule Pune University in India. It can be found for free on Kaggle\textsuperscript{1}. It is a well respected and reviewed dataset. The data consists of 3264 files, which are all images of MRI scans. The images are divided into four sets. There are three different tumor sets, and one set of “no tumor” scans. Considering each image in our dataset is already classified, our model's training is supervised, as opposed to unsupervised, where each image would not have an associated classification to begin with.

\subsection{Preprocessing}

@@

\section{EDA (1 page)}

@@

\begin{figure*}[!ht]
    \centering
    \includegraphics[width=4in]{ImagesPerClassTraining.png}
    \caption{Number of images per class in the training dataset}
    \label{Number of images per class in the training dataset}
\end{figure*}

\begin{figure*}[!ht]
    \centering
    \includegraphics[width=4in]{ImagesPerClassTesting.png}
    \caption{Number of images per class in the testing dataset}
    \label{Number of images per class in the testing dataset}
\end{figure*}

\begin{figure*}[!ht]
    \centering
    \includegraphics[width=4in]{ImageDimensions.png}
    \caption{Dimensions of the images in the dataset}
    \label{Dimensions of images in the dataset}
\end{figure*}

@@

\section{Methodology (1-2 pages)}

@@

\subsection{Early Model Development}

@@

\section{Experimental Results (2 pages)}

\section{Evaluation (1-2 pages)}

\section{Conclusion (1/2 page)}

\section{Our Team (1 page)}
The link to our GitHub repository for this project can be foun at: https://github.com/Coco501/Neural-Networks-for-Neural-Tumors

@@

\subsection{Roles and Contributions}

@@

>>>>

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
% \setlength{\parindent}{1em}
% \setlength{\parskip}{0.8em}
\begin{document}

@@

\author{
\IEEEauthorblockN{\large Joe Vogel}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
jvogel@ucdavis.edu}

\and
\IEEEauthorblockN{\large Nam Nguyen}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
ncnn@ucdavis.edu}

\and
\IEEEauthorblockN{\large Jonathan Mo}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
jtnmo@ucdavis.edu}

\and
\IEEEauthorblockN{\large Aryan Karnwal}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
akarnwal@ucdavis.edu}

\and
\IEEEauthorblockN{\large Richard Ho}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
rdho@ucdavis.edu}
}

@@

\begin{abstract}
This report documents the process of creating a machine learning model that classifies brain tumors using MRI brain scan imagery. This project saw the creation of three successful models using the following algorithms: Convolutional Neural Network, Support Vector Machine, and Random Forest. 
\end{abstract}

\begin{IEEEkeywords}
Brain tumor classification, Convolutional Neural Network (CNN), MRI brain imagery, Random Forest (RF), Support Vector Machine (SVM). 
\end{IEEEkeywords}

\large 
\section{\large Introduction (1-2 pages)}
Hundreds of thousands of MRI brain scans are performed every day. With MRI scans being so intricate, it can sometimes be hard for even a trained professional to determine whether or not a patient has a brain tumor, and even harder to tell which kind.

\subsection{\large Motivation}
Having a tool that could detect and identify tumors within seconds, and with high accuracy, would save a lot of time and money. Not to mention the impact it would have on developing nations who might not have as many specially trained doctors that can readily perform MRI analysis. Early detection is absolutely key to diagnosing and treating these tumors.

\subsection{\large Approach}
A classification machine learning model is the perfect solution to this problem. Using a convolutional neural network and an image dataset, we can train a model to accuractely classify brain tumors. With its complex neural structure, such a model can detect nuances in the scans that are otherwise unnoticeable to the human eye. This project used a dataset that focused on three brain tumor types: glioma, meningioma, pituitary. The model will also be able to predict scans that have "no tumor". Given an image of an MRI brain scan containing one of the four classifications, the model should predict the tumor class with over 95\% accuracy.

\section{\large Literature Review (1/2 page)}

\subsection{\large Brain Tumor Classification in MRI Image Using Convolutional Neural Network$^{1}$}

@@

\subsection{\large A Deep CNN Based Multi-Class Classification of Alzheimer's Disease Using MRI$^{2}$}

@@

\subsection{\large Brain Tumor Classification Using Convolutional Neural Networks$^{3}$}

@@

\section{\large Dataset Description (1 page)}

The dataset used for this project comes from a group of brilliant students at Savitribai Phule Pune University in India. It can be found for free on Kaggle\textsuperscript{4}. It is a well respected and reviewed dataset. The data consists of 3264 files, which are all images of MRI scans. The images are divided into four sets. There are three different tumor sets, and one set of “no tumor” scans. Considering each image in our dataset is already classified, our model's training is supervised, as opposed to unsupervised, where each image would not have an associated classification to begin with.

\subsection{\large Preprocessing}

@@

\section{\large EDA (1 page)}

@@

\begin{figure*}[!ht]
    \centering
    \includegraphics[width=4in]{ImagesPerClassTraining.png}
    \caption{\large Number of images per class in the training dataset}
    \label{Number of images per class in the training dataset}
\end{figure*}

\begin{figure*}[!ht]
    \centering
    \includegraphics[width=4in]{ImagesPerClassTesting.png}
    \caption{\large Number of images per class in the testing dataset}
    \label{Number of images per class in the testing dataset}
\end{figure*}

\begin{figure*}[!ht]
    \centering
    \includegraphics[width=4in]{ImageDimensions.png}
    \caption{\large Dimensions of the images in the dataset}
    \label{Dimensions of images in the dataset}
\end{figure*}

@@

\section{\large Methodology (1-2 pages)}

@@

\subsection{\large Early Model Development}

@@

\section{\large Experimental Results (2 pages)}

\section{\large Evaluation (1-2 pages)}

\section{\large Conclusion (1/2 page)}

\section{\large Our Team (1 page)}
The link to our GitHub repository for this project can be foun at: https://github.com/Coco501/Neural-Networks-for-Neural-Tumors

@@

\subsection{\large Roles and Contributions}

@@

\bibitem{Sartaj2020}
Sartaj. “Brain Tumor Classification (MRI).” Kaggle, 24 May 2020, www.kaggle.com/datasets/sartajbhuvaji/brain-tumor-classification-mri.
